Apache Storm

### 2017-12-10 paper.tex
Base Outline done for submission

\begin{abstract}
In today's world everyone is connected to this ever increasing network of computers be it via a Personal computer, a smart phone or an Embedded Device such as your smart home appliances. These devices generate a huge amounts of data every instance, that can be analyzed to leverage business opportunities previously undiscovered. To be able to discern these windows of opportunities, one must be able to process and analyze the data. In this report we talk about such scenario and the available tools which allows us to process large data streams with minimum latency.
\end{abstract}

@@

\textbf{Data} today is generated by a lot of devices that are connected to the internet and the amount of data being generated at any instance is increasing rapidly. The above mentioned device could be a Personal Computer, a Smart phone, or even an Embedded System. Due to the rapid adoption of Internet of Things, data is being generated at a rate that is unprecedented. The sheer volume of data generated and at the rate at which it is being generated can be termed as Big Data. \textbf{Big Data} made it's appearance in 1998 SGI slide by John Mashey and now is of interest to a lot of financial as well as research institutions.[1]

Data can be considered the oil of the current age, because with huge amounts of data one can determine hidden patterns. A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover something new. The question on how to classify any data as big data can be answered by using the 3Vs of big data: Volume, Variety and Velocity [2]. Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data. In this paper regarding \textbf{Apache Storm} we consider the third V of big data to be the influencing factor.

\textbf{Data Stream} is a sequence of data flowing from a source to a destination, for example a sensor reporting the temperature to a server. The sensor measures the temperature periodically and updates it to the server. Another example could also be the purchase requests of your product. In all of these scenarios, the data is continuously being generated. Traditionally analyzing of these types of data requires storage of these data onto persistence and later batch processing it to discern patterns as well as other useful information. But, with the help of \textbf{Apache Storm} we are able to process these data as they are generated with minimum amount of latency.

This benefit could be a major game changer for many companies as Stream processing of data allows Real Time Analysis. As an example, an online store could monitor its purchase requests for best performing and least performing articles and could offer discounts on the articles not performing good to increase their sales volume. This small change could increase the earnings of a company and will surely be advantageous.

\section{Apache Storm: Concepts}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=.5\textwidth]{arch.png}
    \caption{High Level Storm Architecture}
    \label{fig:arch}
   \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=.5\textwidth]{worker.png}
    \caption{Worker Node Architecture}
    \label{fig:workerarch}
   \end{center}
\end{figure}

@@

\subsubsection{Spout} is the entry point for any data in the Storm Topology. As shown in the figure \ref{fig:topo} spouts take data from external source and emits them into the topology as tuples. The input source could be a Heterogenous data source.

\subsubsection{Bolt} are the processing entities of a Topology. It receives raw tuples emitted by spouts or processed tuples sent by another bolt and sends it down the pipe after it has completed its processing.

@@

\subsection{Practical Implementation}

@@

@@

Ein weiteres, empfehlenswertes Dokument ist das \emph{Kochbuch für \LaTeX{}}, welches auf der Seite \url{http://www.uni-giessen.de/hrz/tex/cookbook/cookbook.html} zu finden ist.

>>>>

\begin{abstract}
In today's world everyone is connected to this ever increasing network of computers be it via a Personal computer, a smart phone or an Embedded Device such as your smart home appliances. These devices generate huge amounts of data every instance, that can be analyzed to uncover patterns, some of which can be used to improve businesses as well. In this report we talk about such scenario and the available tools which allows us to process large data streams with minimum latency. Specifically Apache Storm, a publicly available Stream Processing Tool. We will dive into the concepts of Storm, how one can design usable systems with it and also compare storm to the other available tools.
\end{abstract}

@@

\textbf{Data} today is generated by a lot of devices that are connected to the internet and the amount of data being generated at any instance is increasing rapidly. These devices could include a personal computer, a smart phone, or even an embedded device. Due to the rapid adoption of Internet of Things, data is being generated at an unprecedented rate. The sheer volume of data generated and the rate at which it is being generated can be termed as Big Data. \textbf{Big Data} made it's appearance in 1998 SGI slide by John Mashey and now is of interest to a lot of financial as well as research institutions.[1]

Data can be considered the oil of this present age, because with huge amounts of data one can determine hidden patterns. A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover patterns simply not observable by the human mind(simply due to the volume of data). The question on how to classify any data as big data can be answered by using the 3Vs: Volume, Variety and Velocity [2]. Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data. In this paper regarding \textbf{Apache Storm} we consider the third V of big data to be the influencing factor.

\textbf{Data Stream} is a sequence of data flowing from a source to a destination, for example a sensor reporting the temperature to a server. The sensor measures the temperature periodically and updates it to the server. Another example could also be the purchase requests of products in a shopping site. In all of these scenarios, the data is continuously being generated. Traditionally analyzing of these types of data requires storage of these data onto persistence and later batch processing it to discern patterns as well as other useful information. But, with the help of \textbf{Apache Storm} we are able to process these data as they are generated with minimum amount of latency.

This benefit could be a major game changer for many companies as Stream processing of data allows Real Time Analysis. As an example, an online store could monitor its purchase requests for best performing and least performing articles and could offer discounts on the articles not performing good to increase their sales volume. This small change could increase the gross income of a company and will surely be advantageous.

\section{Apache Storm: Concepts}
\label{sec:concepts}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{arch.png}
    \caption{High Level Storm Architecture}
    \label{fig:arch}
   \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{worker.png}
    \caption{Worker Node Architecture}
    \label{fig:workerarch}
   \end{center}
\end{figure}

@@

\subsubsection{Spout} 
\paragraph{Spouts} are the entry point for any data in the Storm Topology. As shown in the figure \ref{fig:topo} spouts take data from external source and emits them into the topology as tuples. The input source could be a Heterogenous data source.

\subsubsection{Bolt} 
\paragraph{Bolts} are the processing entities of a Topology. It receives raw tuples emitted by spouts or processed tuples sent by another bolt and sends it down the pipe after it has completed its processing.

\section{System Design}
Storm server clusters as discussed above in Section \ref{sec:concepts}, requires at least a Master node and a series of Worker nodes. Apart from these, it also requires zookeeper to keep track of the status of all the master as well as the worker nodes. Setup of these nodes are trivial and can be configured easily in storm using the configuration files. But, the design of a system with regard to the requirements is not such a trivial task because of the different data processing needs. The things to consider while designing a Storm Solution is the System architecture i.e. Lambda or Lambda-less. After selecting one of the above solution, further decisions for the topology design should be considered.

\subsection{System Architecture}
Apache storm is a flexible software which is capable of adapting to the requirements. It can be used in a lambda architecture or a lambda less architecture.Depending upon the requirements of the system, one can opt for either of the above solutions

\subsubsection{Lambda-less Architecture}
\paragraph{Apache Storm} can handle data processing without the need of persistence layer and offers its own 
resiliency mechanisms to make sure data is tuples injected into the system are processed. It has two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, at least once, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure. To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{acker.png}
    \caption{Storm Acker Bolt [4]}
    \label{fig:acker}
   \end{center}
\end{figure}

Figure \ref{fig:acker} shows the Acker bolt integration with the storm topology. It keeps track of the tuple flow starting from the point where it has been injected into the topology till it leaves the topology. After the tuple has left the topology, an \textbf{ack} is sent to the Spout, at this point, it can decide to remove the Tuple from memory.

\subsubsection{Lambda Architecture}
\paragraph{Lambda Architecture} is a data processing Architecture designed to handle high volume data utilizing both batch and stream processing. Apache storm allows users to integrate Persistence models into the system. This flexibility allows users to integrate batch processing into the system as well as seen in figure \ref{fig:lambda}.

Data is persisted in the persistence layer so that in cases of high data volume, Batch processing can be done on the input data to provide guaranteed processing and to reduce the computation requirement on the Real Time Data Processing Unit.

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{lambda.png}
    \caption{Lambda Architecture}
    \label{fig:lambda}
   \end{center}
\end{figure}

\subsection{Topology Design Considerations}
A storm topology design with the spouts and bolts should be properly considered. A spout should inject data into the topology from external sources where as the bolt should be a smallest implementable task. A storm topology should also be considered with various considerations in mind:

\subsubsection{Scalability:} A designed topology should be considered with scalability in mind. The major question is what configuration of bolts enables us to process greater amount of data at real time. Let us consider a simple example of word count topology which counts the repetition of words and ranks the most used words. It can be performed by simply including a spout and two bolts, the first of which to count the number of words and the second to aggregate and rank it (see fig. \ref{fig:ranking} Topology Setup 1).
Since we should have only one instance of Aggregator, to be able to get the aggregated result, in cases of increased data load, the number of rolling count bolts can be increased, but the Aggregation Bolt cannot be increases thus adding increased load at this node. To relax the demand on this node, we can introduce another bolt, the Intermediate Ranking Bolt, which is responsible for ranking the words at a certain predefined intervals and passes the intermediate ranks to the aggregator. This reduces load on the Aggregator since we can increase the intermediate ranking bolts to cope with the increased load.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{ranking.png}
    \caption{Possible Topologies for Word Count and Ranking}
    \label{fig:ranking}
   \end{center}
\end{figure}
\subsubsection{Resilience:} Apache storm includes APIs which allow users to develop solutions which are resilient in nature. Using the acker bolt as show in figure \ref{fig:acker}, users can track the progress of the tuple injected into the system. Thus, writing a spout which can replay the tuple in case of failures allows us to guarantee the processing of data.
\subsubsection{Interoperability:} The designed topology should be interoperable with different sources of data, and by default should work out of the box by adding simply other Spouts in the topology. This also amounts to the maintainability of the code.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{ranking.png}
    \caption{Example Ranking Topologies}
    \label{fig:ranking}
   \end{center}
\end{figure}

@@

\subsection{Practical Implementation}
The implementation of this solution concluded in three sub phases, namely Design, Development and Deployment. In the design phase, the storm topology was thoroughly planned, specifically the spouts, bolts and how they are interconnected. In the development phase, these spouts, bolts and other required aspects of the software was programmed and lastly was deployed on a locally hosted storm server.

\subsubsection{Design}
\paragraph{Designing} a storm topology consists of determining the external sources of data (spouts), the processing that is required to this data imported from external sources (bolts) and the target system. Refer to figure \ref{fig:topoimpl} for the specific topology implementation for our use case. The sources of information for the scope of this paper is only Twitter Streaming API therefore a single Spout for integrating with the Twitter API is sufficient.

To Process the stream of tweets, we need to parse these tweets injected into the topology. The ParseTweetBolt processes these tweets and then emits whether the tweet was positive or negative under the simplified scenario. These separate positive or negative remark on a tweet does not provide useful information unless it is aggregated, thus a ResultAggregationBolt is responsible for aggregating all the results of the Parser bolt and providing the result to the target in real time.

\begin{figure}
  \begin{center}
    \includegraphics[width=.7\textwidth]{topoimpl.png}
    \caption{Topology Implementation for Sentiment Analysis}
    \label{fig:topoimpl}
   \end{center}
\end{figure}

The tweet spout contacts the twitter streaming server and fetches the available tweets in real time and relays it into the topology as a single tweet per tuple. The tuple is forwarded to the TweetParseBolt which will logically analyze the tweet as a positive or a negative tweet and with the percentage probability and emits this result to the ResultAggregationBolt. The ResultAggregationBolt calculates simply the weighted mean of the positive feedback and the negative feedbacks and outputs it into a local file, which can be later observed.

\subsubsection{Development}
\paragraph{Writing} code for storm can be achieved by using different programming languages depending upon the user. Apache Storm was designed to be usable with any programming languages\footnote{\url{http://storm.apache.org/about/multi-language.html}} and currently has Adapters available for Ruby, Python, Javascript and Perl.

The only spout uses Twitter Hosebird Client (\url{https://github.com/twitter/hbc}) to connect to Twitter Streaming Server and consume the Tweets generated in real time. Whereas the Bolts were programmed in java and the code can be viewed in \url{https://github.com/hakku276/storm-sentiment-analysis}.

@@

 <Describe shortly that other tools are available as well and list them with short descriptions>

 \subsection{Comparision between Storm, Spark and Samza}

 <Gather and show the differences between Storm, Spark and Samza in general>

 \subsubsection{Coding}
 <Compare how the coding is in all these platforms>

 \subsubsection{Tool setup}
 <Compare how easy it is to setup the tools>

 \subsubsection{Deployment}
 <Compare how easy it is to deploy the code onto the tool>

 \subsubsection{Resilience}
 <Compare the resilience model of the tools>

 \section{Discussion}
<What I felt about the tool, ease of use and other general aspects>

 The zookeeper services can be run in a single machine or can be deployed in a cluster as well. A single zookeeper instance can also be shared between other services and Apache Storm, but, the major consideration to be taken into account is performance. If there are multiple number of systems that utilize the zookeeper, the Zookeeper can run out of number of clients it can service, thus affecting up-time of the systems in general.[4] Whereas having a dedicated Zookeeper allows a high performing cluster.

@@

Ein weiteres, empfehlenswertes Dokument ist das \emph{Kochbuch für \LaTeX{}}, welches auf der Seite \url{http://www.uni-giessen.de/hrz/tex/cookbook/cookbook.html} zu finden ist.``````````````2
